---
title: "Listener Comfort Scale Preliminary Analysis"
author: "Micah E. Hirsch"
date: "3/7/2022"
output: html_document
---

This is the preliminary analysis rmd file for the listener comfort scale. This analysis is primarily a descriptive analysis to check for trends in the data. Plots/data visualizations will also be generated. This work is part of Peyton Davis' undergraduate Honors in the Major thesis project at Florida State University.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Packages

```{r}

library(rio)
library(tidyverse)
library(furniture)
library(corrr)
library(ggpubr)

```

# Loading the dataset

This is the clean and deidentified version of the dataset. Code showing how this data was prepped can also be found in the following GitHub Repository: https://github.com/micah-hirsch/Preliminary-Listener-Comfort-Scale-Analysis

```{r}

LCscale_data <- rio::import("LCscale_CleanData.csv")

```

# Participant Demographics

These are the demographic information for the participants in our pilot study (age, gender, race, ethnicity, and whether participants were familiar with communication disorders (Familiar_CSD))

```{r}

Listeners <- LCscale_data %>%
  furniture::table1(Age, Gender, Race, Ethnicity, Familiar_CSD)

Listeners

```

# Descriptive Statistics

The descriptive statistics (means and standard deviations) for the average overall scores for the listener comfort scale (LC_Score), Schölderle et al. (2019) overall attitudes scale score (S_Score), and the three subscores from the Schölderle et al. (2019) attitudes scale: Cognition and Language (S_Cogn), Personality (S_personality), and emotional reaction (S_emotion). The first output is for all 10 listeners. The second output groups by the speaker (a healthy control (HC) speaker or a speaker with dysarthria) that the listeners heard during the study.

## Means and SD

```{r}

Descriptives_overall <- LCscale_data %>%
  dplyr::summarize(LC_Score = mean(LC_overallScore), LC_sd = sd(LC_overallScore),
                   S_Score = mean(S_overallScore), S_sd = sd(S_overallScore),
                   S_Cogn = mean(S_cogn_score), S_Cogn_sd = sd(S_cogn_score),
                   S_personality = mean(S_personality_score), S_personality_sd = sd(S_personality_score),
                   S_emotion = mean(S_emotion_score), LC_sd = sd(S_emotion_score))

Descriptives_overall

Descriptives_bySpeaker <- LCscale_data %>%
  dplyr::group_by(Speaker) %>%
  dplyr::summarize(LC_Score = mean(LC_overallScore), LC_sd = sd(LC_overallScore),
                   S_Score = mean(S_overallScore), S_sd = sd(S_overallScore),
                   S_Cogn = mean(S_cogn_score), S_Cogn_sd = sd(S_cogn_score),
                   S_personality = mean(S_personality_score), S_personality_sd = sd(S_personality_score),
                   S_emotion = mean(S_emotion_score), LC_sd = sd(S_emotion_score))

Descriptives_bySpeaker

```

## Data Vis

Bar charts for the overall average scores for the listener comfort scale and the Schölderle et al. (2019) attitudes scale.

```{r}

LC_score_plot <- Descriptives_bySpeaker %>%
  ggplot() +
  aes(x = Speaker,
      y = LC_Score,
      fill = Speaker) +
  geom_col(stat = "identity", fill = c("#782F40","#CEB888")) +
  geom_errorbar(aes(ymin = LC_Score-LC_sd, ymax = LC_Score+LC_sd), width = .2) +
  ylab("Listener Comfort Scale Score") +
  theme_classic()

LC_score_plot

S_score_plot <- Descriptives_bySpeaker %>%
  ggplot() +
  aes(x = Speaker,
      y = S_Score,
      fill = Speaker) +
  geom_col(stat = "identity", fill = c("#782F40","#CEB888")) +
  geom_errorbar(aes(ymin = S_Score-S_sd, ymax = S_Score+S_sd), width = .2) +
  ylab("Listener Attitudes Scale Score") +
  theme_classic()

S_score_plot

Score_plots <- ggarrange(LC_score_plot, S_score_plot, common.legend = T, legend = "bottom", labels = c("A", "B"), nrow = 1, ncol = 2)
ggsave("Plots/Score_plots.png", plot = last_plot(), width = 10, height = 5, units = "in", scale = .8)

Scatterplot <- LCscale_data %>%
  ggplot() +
  aes(x = S_overallScore,
      y = LC_overallScore) +
  geom_point() +
  geom_smooth(method = "lm", se = T, color = "#782F40", fill = "light grey") +
  xlab("Attitudes Rating Overall Scale") +
  ylab("Listener Comfort Overall Score") +
  theme_classic()

ggsave("Plots/LC_Scholderle_scatterplot.png", plot = last_plot(), width = 10, height = 10, units = "in", scale = .8)

```


## Correlations

This a correlation matrix between the overall listener comfort score, the overall attitudes scale (Scholderle et al., 2019) score, and the three subscores from the Schölderle et al. (2019) attitudes scale. One of the main things to note is that the correlation between the listener comfort overall score and the Schölderle attitudes overall score is strong and positive (r=.860, p=.001). 

```{r}

CorrMatrix <- LCscale_data %>%
  dplyr::select(!c(ID:S20)) %>%
  correlate()

CorrMatrix

stats::cor.test(LCscale_data$LC_overallScore, LCscale_data$S_overallScore, method = "pearson")

```

## Group Differences

### LC Score

A t-test was used to test the whether the difference in the overall listener comfort scores are different between the participants who listened to the Dysarthria talker and those who listened to the HC talker. There was no significant different (t(8)=-.5137, p=.6213).

```{r}

# Creating Dysarthria and HC data frames

Dysarthria <- LCscale_data %>%
  dplyr::filter(Speaker == "Dysarthria")

HC <- LCscale_data %>%
  dplyr::filter(Speaker == "HC")

# Assumption Check

## Checking Normality

with(LCscale_data, shapiro.test(LC_overallScore[Speaker == "HC"]))
with(LCscale_data, shapiro.test(LC_overallScore[Speaker == "Dysarthria"]))

## Equal Variance Check

res.ftest.LC <- var.test(LC_overallScore ~ Speaker, data = LCscale_data)
res.ftest.LC

# Model Results

LC_diff <- t.test(HC$LC_overallScore, Dysarthria$LC_overallScore, var.equal = T)
LC_diff

```

### Scholderle Overall Score

A t-test was used to test the whether the difference in the overall Schölderle attitdes scale scores are different between the participants who listened to the Dysarthria talker and those who listened to the HC talker. There was no significant different (t(8)=-1.3578, p=.2116).

```{r}

# Assumption Check

## Checking Normality

with(LCscale_data, shapiro.test(S_overallScore[Speaker == "HC"]))
with(LCscale_data, shapiro.test(S_overallScore[Speaker == "Dysarthria"]))

## Equal Variance Check

res.ftest.S <- var.test(S_overallScore ~ Speaker, data = LCscale_data)
res.ftest.S

# Model Results

S_diff <- t.test(HC$S_overallScore, Dysarthria$S_overallScore, var.equal = T)
S_diff

```

### Scholderle Cognition Subscore

A Mann Whitney U test was used to test the whether the difference in the overall Schölderle attitdes scale scores are different between the participants who listened to the Dysarthria talker and those who listened to the HC talker. (assumptions for the t-test were not met) There was no significant different (U=11, p=.8335).

```{r}

# Assumption Check

## Checking Normality

with(LCscale_data, shapiro.test(S_cogn_score[Speaker == "HC"]))
with(LCscale_data, shapiro.test(S_cogn_score[Speaker == "Dysarthria"]))

## Equal Variance Check

res.ftest.Cogn <- var.test(S_cogn_score ~ Speaker, data = LCscale_data)
res.ftest.Cogn

# Model Results (Mann Whitney U Test)

Cogn_MW <- wilcox.test(S_cogn_score ~ Speaker, data = LCscale_data)
Cogn_MW

```

### Scholderle Personality Subscore

A t-test was used to test the whether the difference in the personality subscore for the Schölderle attitdes scale are different between the participants who listened to the Dysarthria talker and those who listened to the HC talker. There was no significant different (t(8)=-1.423, p=.1925).

```{r}

# Assumption Check

## Checking Normality

with(LCscale_data, shapiro.test(S_personality_score[Speaker == "HC"]))
with(LCscale_data, shapiro.test(S_personality_score[Speaker == "Dysarthria"]))

## Equal Variance Check

res.ftest.Personality <- var.test(S_personality_score ~ Speaker, data = LCscale_data)
res.ftest.Personality

# Model Results

Personality_diff <- t.test(HC$S_personality_score, Dysarthria$S_personality_score, var.equal = T)
Personality_diff

```

### Emotion

A t-test was used to test the whether the difference in the personality subscore for the Schölderle attitdes scale are different between the participants who listened to the Dysarthria talker and those who listened to the HC talker. There was no significant different (t(8)=-1.6493, p=.1377).

```{r}

# Assumption Check

## Checking Normality

with(LCscale_data, shapiro.test(S_emotion_score[Speaker == "HC"]))
with(LCscale_data, shapiro.test(S_emotion_score[Speaker == "Dysarthria"]))

## Equal Variance Check

res.ftest.Emotion <- var.test(S_emotion_score ~ Speaker, data = LCscale_data)
res.ftest.Emotion

# Model Results

Emotion_diff <- t.test(HC$S_emotion_score, Dysarthria$S_emotion_score, var.equal = T)
Emotion_diff

```

# Male and Female Score Comparisons


The two men in our study only rated the healthy control speaker. So this descriptive analysis will only reflect ratings given for the healthy control talker (n=5). 

## Descriptives

For the LC Overall Score

Men: mean = 4.53, SD = 1.30 
Women: mean = 4.65, SD = .69

For the Overall Schölderle Score

Men: mean = 4.03, SD = .60
Women: 4.10, SD = .35

Schölderle mean and SD for subscale scores reported in output below.

```{r}

# Filtering out the listeners who rated the talker with dysarthria. This way we have a dataset with only the ratings for the HC talker.

HC <- LCscale_data %>%
  dplyr::filter(Speaker == "HC")

# Descriptive Statistics

Descriptives_byGender <- HC %>%
  dplyr::group_by(Speaker, Gender) %>%
  dplyr::summarize(LC_Score = mean(LC_overallScore), LC_sd = sd(LC_overallScore),
                   S_Score = mean(S_overallScore), S_sd = sd(S_overallScore),
                   S_Cogn = mean(S_cogn_score), S_Cogn_sd = sd(S_cogn_score),
                   S_personality = mean(S_personality_score), S_personality_sd = sd(S_personality_score),
                   S_emotion = mean(S_emotion_score), LC_sd = sd(S_emotion_score))

Descriptives_byGender

```

# Comparison Based on Familiarity with CSD

Comparing mean and standard deviation of scores for participants familiar with CSD vs. unfamiliar with CSD

## Descriptives

For the LC Overall Score

Dysarthria Talker:
Familiar with CSD: mean = 4.52, SD = .38
Not Familiar with CSD: mean = 5.31, SD = .71

HC Talker:
Familiar with CSD: mean = 5.19, SD = .42
Not Familiar with CSD: mean = 3.72, SD = .94

For the Overall Schölderle Score

Dysarthria Talker:
Familiar with CSD: mean = 4.17, SD = .19
Not Familiar with CSD: mean = 4.90, SD = .57

HC Talker:
Familiar with CSD: mean = 4.35, SD = .10
Not Familiar with CSD: mean = 3.65, SD = .07

Schölderle mean and SD for subscale scores reported in output below.

```{r}

# Descriptive Statistics

Descriptives_byCSD <- LCscale_data %>%
  dplyr::group_by(Speaker, Familiar_CSD) %>%
  dplyr::summarize(LC_Score = mean(LC_overallScore), LC_sd = sd(LC_overallScore),
                   S_Score = mean(S_overallScore), S_sd = sd(S_overallScore),
                   S_Cogn = mean(S_cogn_score), S_Cogn_sd = sd(S_cogn_score),
                   S_personality = mean(S_personality_score), S_personality_sd = sd(S_personality_score),
                   S_emotion = mean(S_emotion_score), LC_sd = sd(S_emotion_score))

Descriptives_byCSD

```

